%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{quickstart}

```{r include=FALSE}
require(knitr)
require(reval)
```

Quickstart for reval
====================

Introduction
------------
This vignette will get you stated using `reval` by showing a moderately-complex 
example. Additional examples are available in the function documentation for
`eval_many`.

Example: Computation of backwater curves
----------------------------------------

In-stream structures such as dams, weirs and culverts modify flows in a river. 
In large rivers with tranquil flows, such structures can affect river flow 
miles upstream. These water surface profiles or "backwater curves" can be 
modelled using well-understood hydraulic models. One important 
parameter---a coefficient representing the material or "roughness" of the river 
bed---is empirical and cannot be easily measured. Therefore it is often 
important for engineers to compute these backwater curves for a range of 
roughness values in order to establish confidence limits for planning and 
management purposes. 

Computing a single profile would look something like this: 

```{r single}
require(rivr)
compute_profile(So = 0.001, n = 0.045, Q = 250, y0 = 2.7, Cm = 1.486, g = 32.2, 
  B = 100, SS = 0, stepdist = 50, totaldist = 3000)
```

If we wanted to loop through a series of roughness values (the argument `n`) 
and compile the results for comparison, we would need a loop:

```{r loop}
ns = seq(0.03, 0.06, by = 0.005)
results = vector("list", length=length(ns))
for(i in 1:length(ns)){
  results[[i]] = compute_profile(So = 0.001, n = ns[i], Q = 250, y0 = 2.7, 
    Cm = 1.486, g = 32.2, B = 100, SS = 0, stepdist = 50, totaldist = 3000)
  # add an identifier to the result
  results[[i]]["n"] = ns[i]
}
combined = do.call(rbind.data.frame, results)
```

That wasn't too bad. But what if we needed to investigate other parameters in 
as well as the roughness coefficient? For each additional variable we would
need an additional for loop, and this would incur a large amount of 
boiler-plate code. `reval` takes care of that stuff for you. We replace the
above loop with a single call to `eval_many`:

```{r reval-single}
results = eval_many(compute_profile, n = seq(0.03, 0.06, by = 0.005),
  default.args = list(So = 0.001, Q = 250, y0 = 2.7, 
  Cm = 1.486, g = 32.2, B = 100, SS = 0, stepdist = 50, totaldist = 3000))
```

Extending the analysis to multiple variables is trivial with `reval`:

```{r reval-multi}
results = eval_many(compute_profile, n = seq(0.03, 0.06, by = 0.005),
  So = c(0.001, 0.0015), default.args = list(Q = 250, 
  y0 = 2.7, Cm = 1.486, g = 32.2, B = 100, SS = 0, stepdist = 50, 
  totaldist = 3000), method = "permute", collate.id = "multi")
```

`reval` adds identifiers automatically, which makes plotting easy.

```{r plot-curves}
require(ggplot2)
ggplot(results, aes(x=x, y=y, color=factor(n))) + geom_line() + 
  facet_wrap(~So)
```

Still confused? Check out the `eval_many` function documentation for more 
examples.
