%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{quickstart}

```{r include=FALSE}
require(knitr)
require(reval)
require(ggplot2)
require(dplyr)
opts_chunk$set(tidy=TRUE, message=FALSE)
```

Quickstart for reval
====================

Introduction
------------
This vignette will get you stated using `reval`. Additional examples are 
available in the `evalmany` function documentation.

Example: Channel design sensitivity analysis
--------------------------------------------

In-stream structures such as dams, weirs and culverts modify flows in a river. 
In large rivers with tranquil flows, such structures can affect the river stage
(water depth) many miles upstream. These water surface profiles or "backwater 
curves" can be modelled using well-understood hydraulic relationships. One 
important parameter---a coefficient representing the texture or "roughness" of 
the river bed---is empirical and cannot be easily measured. Therefore it is 
often important for engineers to compute these backwater curves for a range of
roughness values in order to establish confidence limits for planning and 
management purposes.

The `rivr` package provides the function `compute_profile` for modelling 
backwater curves in prismatic channels given a known water depth at a specified 
location. Computing a single profile would look something like this: 

```{r single}
# install.packages("rivr")
require(rivr)
myprofile = compute_profile(So = 0.001, n = 0.045, Q = 250, y0 = 2.5, 
  Cm = 1.486, g = 32.2, B = 100, SS = 0, stepdist = 50, totaldist = 3000)
head(myprofile)
```

In order to perform a sensitivity analysis on the effect of the roughness 
parameter on the backwater curve, we need to loop through a series of roughness 
values (the argument "n") and compile the results for comparison. This could be 
accomplished with a simple for loop: 

```{r loop}
# loop through values of n
ns = seq(0.03, 0.06, by = 0.005)
results = vector("list", length=length(ns))
for(i in 1:length(ns)){
  results[[i]] = compute_profile(So = 0.001, n = ns[i], Q = 250, y0 = 2.5, 
    Cm = 1.486, g = 32.2, B = 100, SS = 0, stepdist = 50, totaldist = 3000)
  # add an identifier to the result
  results[[i]]["n"] = ns[i]
}
# combine outputs
combined = do.call(rbind.data.frame, results)
```
 
That wasn't too bad, but it's even easier with `reval`. We can replace that
entire chunk of code with a single call to `evalmany`:

```{r reval-single}
results = evalmany(compute_profile, n = seq(0.03, 0.06, by = 0.005),
  default.args = list(So = 0.001, Q = 250, y0 = 2.5, Cm = 1.486, g = 32.2, 
  B = 100, SS = 0, stepdist = 50, totaldist = 3000))
```
```{r plot-single, echo=FALSE, fig.width=10, fig.height=5, dpi=200}
ggplot(results, aes(x = x, y = y, color = reval.set)) + geom_line()
```

Other factors can also influence the water surface profile, such as the channel
slope and the angle of the channel walls. If we wanted to perform a sensitivity
analysis based on these parameters as well, we would need to use a set 
of nested loops or add code to organize the sequence of parameter sets to 
evaluate. This would impact readability and make code maintenance a hassle. 
In contrast, performing a multi-parameter sensitivity analysis with `reval` is 
trivial. Here we use `evalmany` to evaluate all possible permutations of the 
three parameters. When performing more intensive computations, parallel 
processing capabilities from the `doParallel` package can be leveraged using 
the argument `clusters`.

```{r reval-multi}
results = evalmany(compute_profile, n = seq(0.03, 0.06, by = 0.005),
  So = seq(0.001, 0.0015, by = 0.00025), SS = seq(0, 6, by = 2),
  default.args = list(Q = 250, y0 = 2.5, Cm = 1.486, g = 32.2, B = 100, 
  stepdist = 50, totaldist = 3000), method = "permute", collate.id = "multi",
  clusters = 2)
```

The `collate.id` argument is particularly useful for plotting with `ggplot2`
and for subsetting data with `dplyr`.

```{r plot-curves, fig.width=10, dpi=200}
require(ggplot2)
ggplot(results, aes(x = x, y = y, color = factor(n))) + geom_line() + 
  facet_grid(SS ~ So)  
```
```{r filter-data, results="hide"}  
require(dplyr)
filter(results, n == 0.045, So == 0.0015, SS == 2)
```
```{r first-five, echo=FALSE}
kable(head(filter(results, n == 0.045, So == 0.0015, SS == 2)))
```

In this case `compute_profile` outputs a dataframe and collating works as 
expected. When the input function does not produce a dataframe, `evalmany` will
attempt to coerce the output using `as.data.frame`. An optional function to 
further manipulate the raw output prior to coercion to a data frame can be 
passed via the argument `collate.fun`. Alternatively, the raw results can be 
output as a named list with `collate = FALSE`. 


Still confused? 
---------------

Check out the `evalmany` function documentation for more examples.
